---
title: "Final Project | STAT 419"
subtitle: "Group 4"
author: "Abby Drongpa, Alex Sullivan, Chris Liu, Jett Palmer, Sumanth Thokala"
date: 06-08-2025
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

{{< pagebreak >}}

```{r packages}

library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(ggplot2)
library(knitr)

source(here("all_custom_functions.R"))
theme_set(theme_minimal())
```

```{r data}
cincy <- read_csv(here("data", "cincy.csv")) |>
  rename(Pricerank = PRICERANK,
         Age = AGE)
```

## Introduction

Our data set is a subset of home in Cincinnati, Ohio, from the year 2002. It includes 552 observations and seven variables, each representing an individual house. This data set is meant to identify factors that influence and may predict the selling of homes.

Our ordinal grouping variable, "Pricerank", categorizes home based on their selling prices. Houses sold for less than \$190,000 are coded as 1, those sold between \$190,000 and \$285,000 are coded as 2, and houses sold for more than \$285,000 are coded as 3.

The remaining six variables are explanatory:

-   Area: The size of the house in square feet.

-   BR: The number of bedrooms.

-   BA: The number of bathrooms.

-   Garage: The number of cars the garage(s) can accommodate.

-   Quality: An index of construction quality, where 1 indicates high quality, 2 indicates medium quality, and 3 indicates low quality.

-   Age: The age of the house as of 2002.

## Graphs and Summary Statistics

### House Area Histogram

```{r}
cincy |> ggplot(aes(x=Area)) + geom_histogram() + 
  labs(x="Area (Square Feet)", y = "Count", title="House Size")
```

```{r}
summary(cincy$Area)
kable(round(sd(cincy$Area), 3), col.names = "St Dev")
```

The distribution is right skewed with a mean area of 2261 ft^2^ and a median of 2061 ft^2^. The standard deviation is 711.06 ft^2^.

### Car Storage Histogram

```{r}
cincy |> ggplot(aes(x=Garage)) + geom_histogram() + 
  labs(x="Number of Cars", y = "Count", title="Size of Garage")
```

The mean number of cars able to be stored in the garage(s) is 2.1 with a median of 2 cars and a standard deviation of 0.65. The distribution reveals several high and low outliers at 0, 4 , 5, 6, and 7 cars. In other words, the distribution is highly concentrated at 2 car garages.

```{r}
summary(cincy$Garage)
kable(round(sd(cincy$Garage), 3), col.names = "St Dev")
```

### Bedroom Histogram

```{r}
cincy |> ggplot(aes(x=BR)) + geom_histogram() + 
  labs(x="Number of Bedrooms", y = "Count", title="Number of Bedrooms")
```

The distribution of number of bedrooms in each house is roughly normal, with a mean of 3.47 and median of 3 bedrooms. The standard deviation is 1.01 bedrooms.

```{r}
summary(cincy$BR)
kable(round(sd(cincy$BR), 3), col.names = "St Dev")
```

### Bathrooms Histogram

```{r}
cincy |> ggplot(aes(x=BA)) + geom_histogram() + 
  labs(x="Number of Bathrooms", y = "Count", title="Number of Bathrooms")
```

The distribution of number of bathrooms is skewed the right with a mean of 2.64 and median of 3 bedrooms. The standard deviation is 1.06.

```{r}
summary(cincy$BA)
kable(round(sd(cincy$BA), 3), col.names = "St Dev")
```

### Construction Quality Bar Chart

```{r}
cincy |> ggplot(aes(x=Quality)) + geom_histogram() + 
  labs(x="Quality of Construction", y = "Count", title="Construction Quality Distribution")
```

```{r}
summary(cincy$Quality)
kable(round(sd(cincy$Quality), 3), col.names = "St Dev")
```

Most houses were given a quality rating of 2 (medium), followed by 3 (low quality), and 1 (high quality).

## Discriminant and Classification Analysis

### Correlated Quantitative Variables

### Discriminant Analysis

### Classification Analysis

## Summary

## R Code
