---
title: "Final Project | STAT 419"
subtitle: "Group 4"
author: "Abby Drongpa, Alex Sullivan, Chris Liu, Jett Palmer, Sumanth Thokala"
date: 06-08-2025
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

{{< pagebreak >}}

```{r packages}

library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(ggplot2)
library(knitr)

source(here("all_custom_functions.R"))
theme_set(theme_minimal())
```

```{r data}
cincy <- read_csv(here("data", "cincy.csv")) |>
  rename(Pricerank = PRICERANK,
         Age = AGE)
```

## Introduction

Our data set is a subset of home in Cincinnati, Ohio, from the year 2002. It includes 552 observations and seven variables, each representing an individual house. This data set is meant to identify factors that influence and may predict the selling of homes.

Our ordinal grouping variable, "Pricerank", categorizes home based on their selling prices. Houses sold for less than \$190,000 are coded as 1, those sold between \$190,000 and \$285,000 are coded as 2, and houses sold for more than \$285,000 are coded as 3.

The remaining six variables are explanatory:

-   Area: The size of the house in square feet.

-   BR: The number of bedrooms.

-   BA: The number of bathrooms.

-   Garage: The number of cars the garage(s) can accommodate.

-   Quality: An index of construction quality, where 1 indicates high quality, 2 indicates medium quality, and 3 indicates low quality.

-   Age: The age of the house as of 2002.

## Graphs and Summary Statistics

### House Area Histogram

```{r}
cincy |> ggplot(aes(x=Area)) + geom_histogram() + 
  labs(x="Area (Square Feet)", y = "Count", title="House Size")
```

```{r}
summary(cincy$Area)
kable(round(sd(cincy$Area), 3), col.names = "St Dev")
```

The distribution is right skewed with a mean area of 2261 ft^2^ and a median of 2061 ft^2^. The standard deviation is 711.06 ft^2^.

### Car Storage Histogram

```{r}
cincy |> ggplot(aes(x=Garage)) + geom_histogram() + 
  labs(x="Number of Cars", y = "Count", title="Size of Garage")
```

The mean number of cars able to be stored in the garage(s) is 2.1 with a median of 2 cars and a standard deviation of 0.65. The distribution reveals several high and low outliers at 0, 4 , 5, 6, and 7 cars. In other words, the distribution is highly concentrated at 2 car garages.

```{r}
summary(cincy$Garage)
kable(round(sd(cincy$Garage), 3), col.names = "St Dev")
```

### Bedroom Histogram

```{r}
cincy |> ggplot(aes(x=BR)) + geom_histogram() + 
  labs(x="Number of Bedrooms", y = "Count", title="Number of Bedrooms")
```

The distribution of number of bedrooms in each house is roughly normal, with a mean of 3.47 and median of 3 bedrooms. The standard deviation is 1.01 bedrooms.

```{r}
summary(cincy$BR)
kable(round(sd(cincy$BR), 3), col.names = "St Dev")
```

### Bathrooms Histogram

```{r}
cincy |> ggplot(aes(x=BA)) + geom_histogram() + 
  labs(x="Number of Bathrooms", y = "Count", title="Number of Bathrooms")
```

The distribution of number of bathrooms is skewed the right with a mean of 2.64 and median of 3 bedrooms. The standard deviation is 1.06.

```{r}
summary(cincy$BA)
kable(round(sd(cincy$BA), 3), col.names = "St Dev")
```

### Construction Quality Bar Chart

```{r}
cincy |> ggplot(aes(x=Quality)) + geom_histogram() + 
  labs(x="Quality of Construction", y = "Count", title="Construction Quality Distribution")
```

```{r}
summary(cincy$Quality)
kable(round(sd(cincy$Quality), 3), col.names = "St Dev")
```

Most houses were given a quality rating of 2 (medium), followed by 3 (low quality), and 1 (high quality).

## Discriminant and Classification Analysis

### Correlated Quantitative Variables

```{r}
kable(cor(cincy |> select(-Pricerank)))
```

```{r}
plot(cincy[,2:7])
```

We will consider correlations above \|0.80\| to be strong. Provided that no correlations exceed this threshold, we will proceed with all six variables.

### Discriminant Analysis

```{r}
discrim(cincy[,-1], cincy$Pricerank)$a.stand
```

$$
Z_1 = 0.705(y_1) + 0.048(y_2) + 0.330(y_3) + 0.113(y_4) - 0.390(y_5) - 0.223(y_6)
$$

$$
Z_2 = -0.546(y_1) + 0.436(y_2) + 0.086(y_3) - 0.010(y_4) - 0.385(y_5) + 0.070(y_6)
$$

```{r}
kable(data.frame(Variable = c("y1", "y5", "y3", "y6", "y4", "y2"),
                 Z1 = c(0.70478207,
                        -0.38983060,
                        0.33024101,
                        -0.22310228,
                        0.11313689,
                        0.04815262)),
      digits = 3,
      col.names = c("Variable", "Z1"))

kable(data.frame(Variable = c("y1", "y2", "y5", "y3", "y6", "y4"),
                 Z2 = c(-0.54597681,
                        0.43649046,
                        -0.38524224,
                        0.08065737,
                        0.07040947,
                        -0.01035619)),
      digits = 3,
      col.names = c("Variable", "Z2"))

```

The standardized coefficients are ranked above according to their variable contributions to group separation in the presence of all other variables.

H0: $\pmb{\alpha_1} = \pmb{\alpha_2} = \pmb{0}$

Ha: At least one $\pmb{\alpha_i} \neq \pmb{0}$ for i = 1, 2

```{r}
test <- discr.sig(cincy[,-1], cincy$Pricerank)
kable(test, digits = 3, col.names = c("Lambda",
                                      "V",
                                      "P-Values"))
```

<!--I might be getting these hypotheses mixed up with the wrong kind of test. Let's double check together on 6/5 in class. -JP-->

```{r}
cincy$Pricerank <- as.factor(cincy$Pricerank)
discr.plot(cincy[, -1], cincy$Pricerank)
```

The plot of the first two linear discriminant functions shows that LD1 creates meaningful separation between the Pricerank groups. Group 1 (circles) and Group 3 (pluses) appear on opposite ends of the LD1 axis, indicating strong discrimination between lower and higher priced homes. Group 2 (triangles), however, overlaps with both groups, particularly along LD2. This suggests that while the first discriminant function captures most of the between group variation, the second function contributes less to distinguishing the groups. Overall, the functions separate the groups moderately well, but the separation isn’t perfect, especially between groups 2 and 3.

### Classification Analysis

The four most important variables from our discriminant analysis were y1, y2, y5, and y3. We will use these four variables to construct linear classification functions and evaluate classification performance.

```{r}
top_vars <- c("Area", "BR", "Quality", "BA")

lin_out <- lin.class(cincy[, top_vars], cincy$Pricerank)

coefs <- as.data.frame(lin_out$coefs)
coefs$c0 <- lin_out$c.0
rownames(coefs) <- levels(as.factor(cincy$Pricerank))

kable(coefs, digits = 3, caption = "Linear Classification Function Coefficients (Top 4 Variables)")

```

Each group has its own classification function of the form

$$
\text{Score}_k = \pmb{b}_k^\top \pmb{x} + c_{0k}
$$

where ( \pmb{x} ) is the vector of the four selected predictors for an observation, and ( \pmb{b}*k ) and ( c*{0k} ) are the corresponding coefficients for group ( k ).

Classification for Observation #1

```{r}
obs1 <- as.numeric(cincy[1, top_vars])
group_names <- rownames(coefs)

scores <- apply(coefs[, -ncol(coefs)], 1, function(b) sum(b * obs1)) + coefs$c0
predicted <- group_names[which.max(scores)]
actual <- as.character(cincy$Pricerank[1])

cat("Predicted Group:", predicted, "\n")
cat("Actual Group:", actual, "\n")
cat("Correct Prediction:", predicted == actual, "\n")

```

The classification was accurate for observation #1.

Performance Analysis:

```{r}
class_results <- rates(cincy[, top_vars], cincy$Pricerank, method = "l")

kable(class_results$`Confusion Matrix`, caption = "Confusion Matrix")

cat("ACCR:", round(class_results$`Correct Class Rate`, 3), "\n")
cat("AER:", round(class_results$`Error Rate`, 3), "\n")

```

The linear classification functions based on Area, BR, Quality, and BA achieve an apparent correct classification rate of 77%, which indicates a reasonably strong ability to separate the three Pricerank groups. The confusion matrix shows that most misclassifications occur between similar ranking categories, which is expected.

## Summary

## R Code
