---
title: "part_ab"
format: html
editor: visual
---

---
title: "part_ab"
format: html
editor: visual
---

```{r}
cincy <- read.csv("/Users/alexsullivan/Documents/stat419/cincy_419.csv")
library(ggplot2)
library(dplyr)
```

# Section A

Our data set is a subset of homes in Cincinnati, Ohio from 2002 that examined the houses in an effort to determine the factors that affect the selling price. It contains 522 observations and 7 variables. Each observation is corresponds to a house.

One of these variables, “PRICERANK”, is our ordinal grouping variable that classifies homes based on their selling price. Homes sold for less than 190k were labeled as "1", homes sold for between 190k and 285k were labelled with 2, and homes sold for more than 285k were labelled as 3. The other six variables are explanatory. "Area" is measured in square feet and communicates the area of the house. "BR" indicates the number of bedrooms and "BA" indicates the number of bathrooms. "Garage" denotes the number of cars a house can hold. "Quality" is an index measuring the quality of construction where 1 conveys high quality, 2 for medium quality, and 3 for low quality. Finally, "Age" represents the age of the house in 2002.

# Section B

### House Area Histogram

```{r}
cincy |> ggplot(aes(x=Area)) + geom_histogram() + 
  labs(x="Area (Squared Feet)", y = "Count", title="House Area Distribution")
```

```{r}
summary(cincy$Area)
sd(cincy$Area)
```

The mean area of the house is 2261 ft\^2 and the median is 2061 ft\^2. The standard deviation is 711.06. The histogram in Figure 1 is skewed to the right and centered at around 2000 ft\^2.

### Car Storage Histogram

```{r}
cincy |> ggplot(aes(x=Garage)) + geom_histogram() + 
  labs(x="Number of Cars", y = "Count", title="Distribution of number of cars able to be in Garage")
```

The mean number of cars able to be stored in the garage is 2.1 and the median number of cars is 2. The standard deviation is 0.65. This reflects the histogram as the majority of homes store 1-3 cars. There are outliers at 0 cars, 4 cars, 5 cars, 6 cars, and 7 cars. Figure 3 is fairly normal.

```{r}
summary(cincy$Garage)
sd(cincy$Garage)
```

### Bedroom Histogram

```{r}
cincy |> ggplot(aes(x=BR)) + geom_histogram() + 
  labs(x="Number of Bedrooms", y = "Count", title="Number of Bedrooms Distribution")
```

The histogram of the number of bedrooms in each house is pretty normally distributed. It is centered around 3 and this reflects the median of 3 and the mean of 3.47. The standard deviation is 1.01.

```{r}
summary(cincy$BR)
sd(cincy$BR)
```

### Bathrooms Histogram

```{r}
cincy |> ggplot(aes(x=BA)) + geom_histogram() + 
  labs(x="Number of Bathrooms", y = "Count", title="Number of Bathrooms Distribution")
```

The histogram in Figure 5 is slightly skewed to the right. The mean number of bathrooms is 2.64 and the median number of bathrooms is 3. The standard deviation is 1.06.

```{r}
summary(cincy$BA)
sd(cincy$BA)
```

### Construction Quality Bar Chart

```{r}
cincy |> ggplot(aes(x=Quality)) + geom_histogram() + 
  labs(x="Quality of Construction", y = "Count", title="Construction Quality Distribution")
```

```{r}
summary(cincy$Quality)
sd(cincy$Quality)
```
