---
title: "part_e"
format: html
---

# R Code

## Section A

Set up.
```{r, echo=TRUE, results='hide', fig.show='hide', message=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(knitr)
source(here("all_custom_functions.R"))
```

Read in dataframe.
```{r, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
cincy <- read_csv(here("data", "cincy.csv")) |>
  rename(
    Pricerank = PRICERANK,
    Age = AGE
  )
```

## Section B

House area histogram.
```{r, echo=TRUE, results='hide', fig.show='hide', message=FALSE, warning=FALSE}
cincy |> ggplot(aes(x=Area)) + geom_histogram() + 
  labs(x="Area (Squared Feet)", y = "Count", title="House Area Distribution")
```

House area summary statistics.
```{r, echo=TRUE, results='hide'}
summary(cincy$Area)
sd(cincy$Area)
```

Car storage histogram.
```{r, echo=TRUE, results='hide', fig.show='hide', message=FALSE, warning=FALSE}
cincy |> ggplot(aes(x=Garage)) + geom_histogram() + 
  labs(x="Number of Cars", y = "Count", title="Distribution of number of cars able to be in Garage")
```

Car storage summary statistics.
```{r, echo=TRUE, results='hide'}
summary(cincy$Garage)
sd(cincy$Garage)
```

Number of bedrooms histogram.
```{r, echo=TRUE, results='hide', fig.show='hide', message=FALSE, warning=FALSE}
cincy |> ggplot(aes(x=BR)) + geom_histogram() + 
  labs(x="Number of Bedrooms", y = "Count", title="Number of Bedrooms Distribution")
```

Number of bedrooms summary statistics.
```{r, echo=TRUE, results='hide'}
summary(cincy$BR)
sd(cincy$BR)
```

Number of bathrooms histogram.
```{r, echo=TRUE, results='hide', fig.show='hide'}
cincy |> ggplot(aes(x=BA)) + geom_histogram() + 
  labs(x="Number of Bathrooms", y = "Count", title="Number of Bathrooms Distribution")
```

Number of bathrooms summary statistics.
```{r, echo=TRUE, results='hide'}
summary(cincy$BA)
sd(cincy$BA)
```

Construction quality bar chart.
```{r, echo=TRUE, results='hide', fig.show='hide', , message=FALSE, warning=FALSE}
cincy |> ggplot(aes(x=Quality)) + geom_histogram() + 
  labs(x="Quality of Construction", y = "Count", title="Construction Quality Distribution")
```

Construction quality summary statistics.
```{r, echo=TRUE, results='hide'}
summary(cincy$Quality)
sd(cincy$Quality)
```

Correlation coefficients between all quantitative explanatory variables.
```{r, echo=TRUE, results='hide'}
kable(cor(cincy |> select(-Pricerank)))
```

Plot of correlation between all quantitative explanatory variables.
```{r, echo=TRUE, results='hide', fig.show='hide'}
plot(cincy[,2:7])
```

Discriminant analysis, standardised functions.
```{r, echo=TRUE, results='hide'}
discrim(cincy[,-1], cincy$Pricerank)$a.stand
```

Ordered standardised coefficients for both functions, by absolute value from zero, descending.
```{r, echo=TRUE, results='hide'}
kable(data.frame(Variable = c("y1", "y5", "y3", "y6", "y4", "y2"),
                 Z1 = c(0.70478207,
                        -0.38983060,
                        0.33024101,
                        -0.22310228,
                        0.11313689,
                        0.04815262)),
      digits = 3,
      col.names = c("Variable", "Z1"))

kable(data.frame(Variable = c("y1", "y2", "y5", "y3", "y6", "y4"),
                 Z2 = c(-0.54597681,
                        0.43649046,
                        -0.38524224,
                        0.08065737,
                        0.07040947,
                        -0.01035619)),
      digits = 3,
      col.names = c("Variable", "Z2"))
```

Hypothesis test on discriminant functions.
```{r, echo=TRUE, results='hide'}
test <- discr.sig(cincy[,-1], cincy$Pricerank)
kable(test, digits = 3, col.names = c("Lambda",
                                      "V",
                                      "P-Values"))
```

Plot of discriminant functions.
```{r, echo=TRUE, results='hide', fig.show='hide'}
cincy$Pricerank <- as.factor(cincy$Pricerank)
discr.plot(cincy[, -1], cincy$Pricerank)
```


Classification analysis with top four variables.
```{r, echo=TRUE, results='hide', , message=FALSE, warning=FALSE}
top_vars <- c("Area", "BR", "Quality", "BA")

lin_out <- lin.class(cincy[, top_vars], cincy$Pricerank)

coefs <- as.data.frame(lin_out$coefs)
coefs$c0 <- lin_out$c.0
rownames(coefs) <- levels(as.factor(cincy$Pricerank))

kable(coefs, digits = 3, caption = "Linear Classification Function Coefficients (Top 4 Variables)")
```

Classification for the first observation (home) in the dataset.
```{r, echo=TRUE, results='hide'}
obs1 <- as.numeric(cincy[1, top_vars])
group_names <- rownames(coefs)

scores <- apply(coefs[, -ncol(coefs)], 1, function(b) sum(b * obs1)) + coefs$c0
predicted <- group_names[which.max(scores)]
actual <- as.character(cincy$Pricerank[1])

cat("Predicted Group:", predicted, "\n")
cat("Actual Group:", actual, "\n")
cat("Correct Prediction:", predicted == actual, "\n")
```

Performance analysis of classification.
```{r, echo=TRUE, results='hide'}
class_results <- rates(cincy[, top_vars], cincy$Pricerank, method = "l")

kable(class_results$`Confusion Matrix`, caption = "Confusion Matrix")

cat("ACCR:", round(class_results$`Correct Class Rate`, 3), "\n")
cat("AER:", round(class_results$`Error Rate`, 3), "\n")
```

