---
title: "Stat 419 Final Project"
author: "Sumanth, Abby, Jett, Alex, Chris"
format: html
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

```{r packages}
# Echo TRUE for this whole file. The final document will be echo FALSE, then include code at the bottom (section E)
library(tidyverse)
library(here)
library(janitor)
library(knitr)
source(here("all_custom_functions.R"))
```

```{r data}
cincy <- read_csv(here("data", "cincy.csv")) |>
  rename(
    Pricerank = PRICERANK,
    Age = AGE
  )
```

## Discriminant and Classification Analysis

### Correlated Quantitative Variables

```{r}
kable(cor(cincy |> select(-Pricerank)))
```

```{r}
plot(cincy[,2:7])
```

We will consider correlations above \|0.80\| to be strong. Provided that no correlations exceed this threshold, we will proceed with all six variables.

### Discriminant Analysis

```{r}
discrim(cincy[,-1], cincy$Pricerank)$a.stand
```

$$
Z_1 = 0.705(y_1) + 0.048(y_2) + 0.330(y_3) + 0.113(y_4) - 0.390(y_5) - 0.223(y_6)
$$

$$
Z_2 = -0.546(y_1) + 0.436(y_2) + 0.086(y_3) - 0.010(y_4) - 0.385(y_5) + 0.070(y_6)
$$

```{r}
kable(data.frame(Variable = c("y1", "y5", "y3", "y6", "y4", "y2"),
                 Z1 = c(0.70478207,
                        -0.38983060,
                        0.33024101,
                        -0.22310228,
                        0.11313689,
                        0.04815262)),
      digits = 3,
      col.names = c("Variable", "Z1"))

kable(data.frame(Variable = c("y1", "y2", "y5", "y3", "y6", "y4"),
                 Z2 = c(-0.54597681,
                        0.43649046,
                        -0.38524224,
                        0.08065737,
                        0.07040947,
                        -0.01035619)),
      digits = 3,
      col.names = c("Variable", "Z2"))

```

The standardized coefficients are ranked above according to their variable contributions to group separation in the presence of all other variables.

H0: $\pmb{\alpha_1} = \pmb{\alpha_2} = \pmb{0}$

HA: At least one $\pmb{\alpha_i} \neq \pmb{0}$ for i = 1, 2

```{r}
test <- discr.sig(cincy[,-1], cincy$Pricerank)
kable(test, digits = 3, col.names = c("Lambda",
                                      "V",
                                      "P-Values"))
```

<!--I might be getting these hypotheses mixed up with the wrong kind of test. Let's double check together on 6/5 in class. -JP-->

### Classification Analysis

The four most important variables from our discriminant analysis were y1, y2, y5, and y3. We will use these four variables to construct linear classification functions and evaluate classification performance.

```{r}
top_vars <- c("Area", "BR", "Quality", "BA")

lin_out <- lin.class(cincy[, top_vars], cincy$Pricerank)

coefs <- as.data.frame(lin_out$coefs)
coefs$c0 <- lin_out$c.0
rownames(coefs) <- levels(as.factor(cincy$Pricerank))

kable(coefs, digits = 3, caption = "Linear Classification Function Coefficients (Top 4 Variables)")

```

Each group has its own classification function of the form

$$
\text{Score}_k = \mathbf{b}_k^\top \mathbf{x} + c_{0k}
$$

where \( \mathbf{x} \) is the vector of the four selected predictors for an observation, and \( \mathbf{b}_k \) and \( c_{0k} \) are the corresponding coefficients for group \( k \).


Classification for Observation #1

```{r}
obs1 <- as.numeric(cincy[1, top_vars])
group_names <- rownames(coefs)

scores <- apply(coefs[, -ncol(coefs)], 1, function(b) sum(b * obs1)) + coefs$c0
predicted <- group_names[which.max(scores)]
actual <- as.character(cincy$Pricerank[1])

cat("Predicted Group:", predicted, "\n")
cat("Actual Group:", actual, "\n")
cat("Correct Prediction:", predicted == actual, "\n")

```

The classification was accurate for observation #1

Performance Analysis:

```{r}
class_results <- rates(cincy[, top_vars], cincy$Pricerank, method = "l")

kable(class_results$`Confusion Matrix`, caption = "Confusion Matrix")

cat("ACCR:", round(class_results$`Correct Class Rate`, 3), "\n")
cat("AER:", round(class_results$`Error Rate`, 3), "\n")

```
The linear classification functions based on Area, BR, Quality, and BA achieve an apparent correct classification rate of 77%, which indicates a reasonably strong ability to separate the three Pricerank groups. The confusion matrix shows that most misclassifications occur between similar ranking categories, which is expected.


